---
title: "MAC study"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Setup -------------------------------------------------------------------

# Required packages
pacs <- c("tidyverse", "haven", "tables", "lme4", "lmerTest", "emmeans")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, name) {
  data.frame(num=grep(name, names(df), ignore.case=TRUE),
             var=grep(name, names(df), ignore.case=TRUE, value=TRUE))
}

# Path to data file
setwd("C:\\Users\\keiji\\Dropbox\\Nutrition\\Julie Jones")

# Read data ---------------------------------------------------------------

# Read data, n obs = 70
# mac0 <- read_sav("MAC Endpoint data with baseline values.sav")
mac0 <- read_sav("MAC Endpoint data with baseline values 102121.sav")

# Fix variable names
fixthis <- mac0 %>% search_var("8iso") %>% pull(num)
names(mac0)[fixthis] <- c("isopgf2pgmL", "isopgf2pgmLB")

# Data including all subjects
mac <- mac0 %>% 
  mutate(ID = factor(ID),
         Treatment = factor(Treatment, labels = c("Control", "Mac")),
         Treatment = relevel(Treatment, ref = "Mac"),
         Gender = factor(Gender),
         Group = factor(Group, labels = c("Mac-Control", "Control-Mac")),
         Phase = factor(Phase),
         BaseBMI = ifelse(BMIB >= 30, 1, 0),
         BaseBMI = factor(BaseBMI, labels = c("<30", ">=30")),
         BaseWC = ifelse(WCbaseline >= 108, 1, 0),
         BaseWC = factor(BaseWC, labels = c("<108", ">=108")),
         BaseBF = ifelse(PBFPercentBodyFat >= 43, 1, 0),
         BaseBF = factor(BaseBF, labels = c("<43", ">=43")))

# Female only (for sub-group analysis)
# Changing cut-off values for WC and body fat
mac_fem <- mac %>% 
  filter(Gender == "F") %>% 
  mutate(BaseWC = ifelse(WCbaseline >= 100, 1, 0),
         BaseWC = factor(BaseWC, labels = c("<100", ">=100")),
         BaseBF = ifelse(PBFPercentBodyFat >= 44, 1, 0),
         BaseBF = factor(BaseBF, labels = c("<44", ">=44")))

```

##  Datasets

* File name: `MAC Endpoint data with baseline values 102121.sav`
* Data are from a 2 x 2 crossover design (AB/BA) with control/mac treatments
* Includes *n* = `r nrow(mac0)` observations from `r n_distinct(mac0$ID)` subjects
* There are `r ncol(mac0)` variables including:
  * `ID`
  * Study design variables:
    * Sequence: `Group`
    * `Phase`  
    * `Visit`
    * `Treatment`
  * Demographics & body measurements
    * `Age`
    * `Height`
    * `Weight`
    * `BMI`
    * Waist circumference: `WC`
  * Blood lipids
  * Insulin resistance
  * Inflammatory/oxidative markers

## Descriptive analysis on inflammatory/oxidative markers

* Variables included:
  * `CRPmgdL`
  * `ESelectinngdL`
  * `IL6pgmL`
  * `TNFapgmL`
  * `sICAM1pgmL`
  * `sVCAm1pgmL`
  * `isopgf2pgmL`
  * `MDAnmolmL`

### Mean/SD by treatment

```{r echo = FALSE}
# Variable names
inflam_vars <- c("crp", "eselec", "il6", "tnf", "icam", "vcam", "iso", "mda") %>% 
  map(function(x) grep(x, names(mac0), value = TRUE, ignore.case = TRUE)) %>% 
  sapply("[[", 1)

inflamB_vars <- paste0(inflam_vars, "B")

# Descriptive means
# Data including baseline values as obs
mac_with_base <- mac %>%
  filter(Phase == 1) %>% 
  mutate(Treatment = "Baseline") %>% 
  select(ID, Treatment, all_of(inflamB_vars)) %>% 
  rename_with(function(x) substr(x, 1, nchar(x) - 1), all_of(inflamB_vars)) %>% 
  bind_rows(select(mac, ID, Treatment, all_of(inflam_vars)))

# Mean (SD) by treatment
options(pillar.sigfig = 4)
mac_with_base %>% 
  pivot_longer(any_of(inflam_vars), names_to = "var", values_to = "value") %>% 
  mutate(var = factor(var, levels = inflam_vars)) %>% 
  group_by(var, Treatment) %>% 
  summarize(mean = mean(value), sd = sd(value)) %>% 
  mutate_at(vars(mean, sd), round, 2) %>% 
  knitr::kable()
```

### Mean/SD by sequence group and treatment


```{r echo = FALSE}
# Mean (SD) by sequence and treatment
Mean <- function(x) mean(x, na.rm = TRUE)
SD <- function(x) sd(x, na.rm = TRUE)
tabular((CRPmgdL + ESelectinngdL + IL6pgmL + TNFapgmL + sICAM1pgmL + sVCAm1pgmL + isopgf2pgmL + MDAnmolmL) * (Mean + SD) * Format(digits = 2)  ~  
          Heading() * Group * Heading() * Treatment, data = mac) %>% 
  suppressWarnings()
```
