---
title: "MAC study"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Setup -------------------------------------------------------------------

# Required packages
pacs <- c("tidyverse", "haven", "tables", "lme4", "lmerTest", "emmeans")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, name) {
  data.frame(num=grep(name, names(df), ignore.case=TRUE),
             var=grep(name, names(df), ignore.case=TRUE, value=TRUE))
}

# Path to data file
setwd("C:\\Users\\keiji\\Dropbox\\Nutrition\\Julie Jones")

# Read data ---------------------------------------------------------------

# Read data, n obs = 70
# mac0 <- read_sav("MAC Endpoint data with baseline values.sav")
mac0 <- read_sav("MAC Endpoint data with baseline values 102121.sav")

# Fix variable names
fixthis <- mac0 %>% search_var("8iso") %>% pull(num)
names(mac0)[fixthis] <- c("isopgf2pgmL", "isopgf2pgmLB")

# Data including all subjects
mac <- mac0 %>% 
  mutate(ID = factor(ID),
         Treatment = factor(Treatment, labels = c("Control", "Mac")),
         Treatment = relevel(Treatment, ref = "Mac"),
         Gender = factor(Gender),
         Group = factor(Group, labels = c("Mac-Control", "Control-Mac")),
         Phase = factor(Phase),
         BaseBMI = ifelse(BMIB >= 30, 1, 0),
         BaseBMI = factor(BaseBMI, labels = c("<30", ">=30")),
         BaseWC = ifelse(WCbaseline >= 108, 1, 0),
         BaseWC = factor(BaseWC, labels = c("<108", ">=108")),
         BaseBF = ifelse(PBFPercentBodyFat >= 43, 1, 0),
         BaseBF = factor(BaseBF, labels = c("<43", ">=43")))

# Female only (for sub-group analysis)
# Changing cut-off values for WC and body fat
mac_fem <- mac %>% 
  filter(Gender == "F") %>% 
  mutate(BaseWC = ifelse(WCbaseline >= 100, 1, 0),
         BaseWC = factor(BaseWC, labels = c("<100", ">=100")),
         BaseBF = ifelse(PBFPercentBodyFat >= 44, 1, 0),
         BaseBF = factor(BaseBF, labels = c("<44", ">=44")))

```

##  Datasets

* File name: `MAC Endpoint data with baseline values 102121.sav`
* Data are from a 2 x 2 crossover design (AB/BA) with control/mac treatments
* Includes *n* = `r nrow(mac0)` observations from `r n_distinct(mac0$ID)` subjects
* There are `r ncol(mac0)` variables including:
  * `ID`
  * Study design variables:
    * Sequence: `Group`
    * `Phase`  
    * `Visit`
    * `Treatment`
  * Demographics & body measurements
    * `Age`
    * `Height`
    * `Weight`
    * `BMI`
    * Waist circumference: `WC`
  * Blood lipids
  * Insulin resistance
  * Inflammatory/oxidative markers

## Descriptive analysis on inflammatory/oxidative markers

* Variables included:
  * `CRPmgdL`
  * `ESelectinngdL`
  * `IL6pgmL`
  * `TNFapgmL`
  * `sICAM1pgmL`
  * `sVCAm1pgmL`
  * `isopgf2pgmL`
  * `MDAnmolmL`

### Mean/SD by treatment

```{r echo = FALSE, message = FALSE, warning = FALSE }
# Variable names
inflam_vars <- c("crp", "eselec", "il6", "tnf", "icam", "vcam", "iso", "mda") %>% 
  map(function(x) grep(x, names(mac0), value = TRUE, ignore.case = TRUE)) %>% 
  sapply("[[", 1)

inflamB_vars <- paste0(inflam_vars, "B")

# Descriptive means
# Data including baseline values as obs
mac_with_base <- mac %>%
  filter(Phase == 1) %>% 
  mutate(Treatment = "Baseline") %>% 
  select(ID, Treatment, all_of(inflamB_vars)) %>% 
  rename_with(function(x) substr(x, 1, nchar(x) - 1), all_of(inflamB_vars)) %>% 
  bind_rows(select(mac, ID, Treatment, all_of(inflam_vars))) %>% 
  mutate(Treatment = factor(Treatment))

# Mean (SD) by treatment
Mean <- function(x) mean(x, na.rm = TRUE)
SD <- function(x) sd(x, na.rm = TRUE)
tabular((CRPmgdL + ESelectinngdL + IL6pgmL + TNFapgmL + sICAM1pgmL + sVCAm1pgmL + isopgf2pgmL + MDAnmolmL) * (Mean + SD) * Format(digits = 2)  ~  
          Heading() * Treatment, data = mac_with_base) %>% 
  suppressWarnings()
```

### Mean/SD by sequence group and treatment


```{r echo = FALSE}
# Mean (SD) by sequence and treatment
tabular((CRPmgdL + ESelectinngdL + IL6pgmL + TNFapgmL + sICAM1pgmL + sVCAm1pgmL + isopgf2pgmL + MDAnmolmL) * (Mean + SD) * Format(digits = 2)  ~  
          Heading() * Group * Heading() * Treatment, data = mac) %>% 
  suppressWarnings()
```

## Mixed model analysis on on inflammatory/oxidative markers

* Each of inflammatory/oxidative variables was log-transformed and used as a dependent variable in the following mixed model.
* The mixed model included:
  * treatment(mac/control), sequence groups (AB/BA), and phase as fixed-effects
  * subjects as random-effects
* Marginal means were estimated for both treatments (with 95% CI) adjusted for sequence and phase, and then back-transformed into the original scale. 
* The last row of `Mac - Control` actually refers to the **ratio** of treatment means (mac/control) after the back-transformation.
  * None of these variables showed a significant difference between the two treatments.

```{r echo = FALSE}
tidy_output <- function(emm){
  
  delta <- confint(pairs(emm)) %>% 
    rename(Treatment = contrast, emmean = estimate)
  
  pval <- emm %>% pairs() %>% summary() %>% pull(p.value) %>% 
    finalfit::p_tidy(digits = 4, prefix = NULL) %>% 
    c("", "", .) %>% 
    as_tibble_col(column_name = "pval")
  
  as.data.frame(emm) %>% 
    bind_rows(delta) %>% 
    select(-df, -SE) %>% 
    bind_cols(pval)
}

show_result <- function(emm, log = FALSE){
  emm <- emm %>% tidy_output()
  if(log) emm <- emm %>% mutate_at(2:4, exp)
  emm <- emm %>% 
    mutate_at(2:4, round, 2)
}

# Base model: adjusts for sequence and phase
write_model <- function(y, trt_var = "Treatment") as.formula(paste0(y, " ~ ", trt_var, "+ Group + Phase + (1|ID)"))

# All together
inflam_results <- function(var){
  write_model(y = var) %>% 
    lmer(data = mac) %>% 
    suppressWarnings() %>% 
    emmeans(~ Treatment) %>% 
    show_result(log = TRUE)
}

yvars <- c("log(CRPmgdL)", "log(ESelectinngdL)", "log(IL6pgmL)", "log(TNFapgmL)", 
           "log(sICAM1pgmL)", "log(sVCAm1pgmL)", "log(isopgf2pgmL)", "log(MDAnmolmL)")
all_inflam_results <- yvars %>% map(inflam_results)
names(all_inflam_results) <- yvars
print(all_inflam_results, row.names = FALSE)
```

### Examining the interaction with baseline BMI

* To examine if the effect of mac treatment may be different depending on baseline BMI, a dichotomous BMI variable (<30 and >=30) was added into the mixed model, along with its interaction with treatment.
* Marginal means were estimated for each combination of treatment and baseline BMI adjusted for sequence and phase, and then back-transformed into the original scale. 
* The last row of `Mac - Control` actually refers to the **ratio** of treatment means (mac/control) after the back-transformation.
  * None of these variables showed a significant interaction with baseline BMI, indicating that the effect of mac treatment is not significantly different between BMI <30 and >=30.

```{r echo = FALSE}
tidy_output2 <- function(emm, lmer){
  
  flevels <- summary(emm) %>% select(2) %>% distinct()
  
  pval <- summary(pairs(emm, adjust = "none")) %>% 
    pull(p.value) %>% 
    finalfit::p_tidy(digits = 4, prefix = NULL) %>% 
    as_tibble_col(column_name = "pval")
  
  intxp <- coef(summary(lmer)) %>% as_tibble() %>%
    slice(n()) %>% pull(-1) %>% 
    finalfit::p_tidy(digits = 4, prefix = NULL) %>% 
    c(rep("", 5), .) %>% 
    as_tibble_col(column_name = "intx.P")
  
  delta <- confint(pairs(emm, adjust = "none")) %>% 
    mutate(Treatment = "Mac - Ctrl") %>% 
    rename(emmean = estimate) %>% 
    bind_cols(pval) %>% slice(1, 6) %>% 
    bind_cols(flevels)
  
  summary(emm) %>% 
    mutate(pval = "") %>% 
    bind_rows(delta) %>% 
    select(-contrast, -df, -SE) %>% 
    arrange(.[2]) %>%
    bind_cols(intxp)
}

show_result2 <- function(emm, lmer, log = FALSE){
  emm <- emm %>% tidy_output2(lmer)
  if(log) emm <- emm %>% mutate_at(3:5, exp)
  emm %>% 
    mutate_at(3:5, round, 2) 
    # print(row.names = FALSE)
}

inflam_results <- function(var){
  mod <- write_model(y = var, "Treatment * BaseBMI") %>% 
    lmer(data = mac) %>% 
    suppressWarnings()
  mod %>% 
    emmeans(~ Treatment + BaseBMI) %>% 
    show_result2(mod, log = TRUE)
}

all_inflam_results2 <- yvars %>% map(inflam_results)
names(all_inflam_results2) <- yvars
print(all_inflam_results2, row.names = FALSE)
```

### Examining the interaction with baseline WC

* To examine if the effect of mac treatment may be different depending on baseline waist circumference (WC), a dichotomous WC variable (<108 and >=108) was added into the mixed model, along with its interaction with treatment.
* Marginal means were estimated for each combination of treatment and baseline WC adjusted for sequence and phase, and then back-transformed into the original scale. 
* The last row of `Mac - Control` actually refers to the **ratio** of treatment means (mac/control) after the back-transformation.
  * None of these variables showed a significant interaction with baseline WC, indicating that the effect of mac treatment is not significantly different between WC <108 and >=108.

```{r echo = FALSE}
# Interaction with baseline WC, dichotomous

inflam_results <- function(var){
  mod <- write_model(y = var, "Treatment * BaseWC") %>% 
    lmer(data = mac) %>% 
    suppressWarnings()
  mod %>% 
    emmeans(~ Treatment + BaseWC) %>% 
    show_result2(mod, log = TRUE)
}
all_inflam_results3 <- yvars %>% map(inflam_results)
names(all_inflam_results3) <- yvars
print(all_inflam_results3, row.names = FALSE)
```

### Examining the interaction with baseline % body fat

* To examine if the effect of mac treatment may be different depending on baseline % body fat (BF), a dichotomous BF variable (<43 and >=43) was added into the mixed model, along with its interaction with treatment.
* Marginal means were estimated for each combination of treatment and baseline BF adjusted for sequence and phase, and then back-transformed into the original scale. 
* The last row of `Mac - Control` actually refers to the **ratio** of treatment means (mac/control) after the back-transformation.
  * None of these variables showed a significant interaction with baseline BF, indicating that the effect of mac treatment is not significantly different between WC <43 and >=43.

```{r echo = FALSE}
# Interaction with baseline WC, dichotomous

inflam_results <- function(var){
  mod <- write_model(y = var, "Treatment * BaseBF") %>% 
    lmer(data = mac) %>% 
    suppressWarnings()
  mod %>% 
    emmeans(~ Treatment + BaseBF) %>% 
    show_result2(mod, log = TRUE)
}
all_inflam_results4 <- yvars %>% map(inflam_results)
names(all_inflam_results4) <- yvars
print(all_inflam_results4, row.names = FALSE)
```
